Current and future constraints on primordial magnetic fields

### 2016-10-31 pmf.tex
a

\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\planck}{{\sl Planck}}
\newcommand{\wmap}{{\sl WMAP}}
\newcommand{\bicepkeck}{BICEP2/Keck}
\newcommand{\sptnew}{SPT-3G}
\newcommand{\pb}{POLARBEAR}
\newcommand{\simons}{Simons Array}
\newcommand{\sptpol}{SPTpol}
\newcommand{\advactpol}{Adv.~ACTpol}

@@

We present new limits on the amplitude of potential primordial magnetic fields (PMFs) using temperature and polarization measurements of the cosmic microwave background (CMB)  from \planck{}, \bicepkeck{}, \pb, and \sptpol. 
We reduce twofold the upper limit on the CMB anisotropy power due to a PMF, from $\apmf < 0.73$ from Planck alone to $\apmf < 0.35$ for the combined dataset at 95\% CL. 
%Most of the improvement is due to the addition of the \bicepkeck{} data; without these bandpowers, the combined limit weakens to $\apmf < 0.\tbd{XX}$. 
We also forecast the expected future limits from stage III CMB experiments (like \sptnew{},  \advactpol, or the \simons) and the proposed CMB stage IV experiment. 
Future CMB experiments should dramatically reduce the current uncertainties, by two orders of magnitude for stage III experiments and three orders of magnitude for stage IV. 
The constraints from a stage IV experiment have the potential to rule out much of the parameter space for PMFs.

@@

CMB B-mode measurements can also be used to constrain more exotic models, such as the possible existence of cosmic birefringence \citep{carroll98,lue99} or a primordial magnetic field (PMF) \citep{kosowsky96, seshadri01}.  
Both effects lead to cosmic birefringence and the rotation of E-modes into B-modes. 
By equating the magnitude of the resulting B-modes, parity-violating processes can be translated into an equivalent PMF strength. 
We will therefor simply quote effective limits on a PMF. 
In this work, we present new upper limits from current CMB polarization data on the possibility of a PMF or parity-violating physics.

@@

Primordial magnetic fields would have observational consequences for Big Bang nucleosynthesis \citep[e.g.,][]{kahniashvili10}, large scale structure \citep[e.g.,][]{battaner97}, ionization history of the Universe \citep{kunze15}, and the black body spectrum of the CMB \citep[e.g.,][]{kunze14},  as well as the CMB anisotropy. 
The CMB anisotropies have yielded some of the strongest constraints on PMFs and are the focus of this work.
There have been two recent results of note. 
\citet{planck15-19} have used the \planck{} 2015 release of temperature and polarization data to set limits on a variety of PMF models. 
With the CMB power spectrum data that will be the focus of this paper, \citet{planck15-19} find 95\% CL upper limits ranging from $\bpmf < 5.6$\,nG to $<0.7$\,nG depending on the exact model. 
The \pb{} collaboration also recently announced limits on PMFs from the \pb{} data using either a 4-point estimator or B-mode power spectrum measurement in \citep{polarbear14b}. 
The strongest constraints were from the B-mode spectrum; the observed upper limit was $\bpmf < 3.9$\,nG. 
Two other experiments, \sptpol{} and \bicepkeck{}, have reported B-mode power spectrum measurements recently \citep{keisler15,bicepkeck15}.
In this paper, we will combine the data from all four experiments to determine simple limits on the PMF, and then examine Fisher matrix forecasts on PMF models from the stage-III and stage-IV CMB experiments being built or designed right now.

@@

In addition to the \planck{} data, we use a number of recent CMB polarization measurements. 
First, we include TE and EE spectrum measurements\citep{crites15} for $\ell \in [500,3000]$ and BB bandpowers covering $\ell \in [500,2000]$ \citep{keisler15} from SPTpol. 
We also add the BB bandpowers from \pb{} that cover the multipoles from 500 to 2500 \citep{polarbear14b}. 
Both the \sptpol{} and \pb{} bandpowers primary constrain the vector mode of the PMF given the angular scales measured. 
Finally, we include the latest BICEP2 and Keck Array  joint analysis \citep{bicepkeck15}. 
This latter dataset also places limits on the tensor mode of the PMF due to its coverage of lower multipoles. 
During the writing of this work, ACTpol polarization power spectra became available \citep{naess14,louis16}. 
We have not included them as the public likelihoods do not include the B-mode power spectrum; we do not expect adding the ACTpol bandpowers to significantly change the PMF limits based on a visual comparison of the current BB bandpowers from different experiments.
In all chains, we marginalize over the recommended foreground models for each data set. 
We do not however require these consistency between these foreground models since the data don't have identical flux cuts and so on. 
In the case of an upper limit this treatment is conservative since wider foreground priors should only weaken the limit on the PMF strength.

@@

We use MCMC methods to determine the parameter constraints reported in this work. 
The results are calculated using  the {\textsc CosmoMC}\footnote{http://cosmologist.info/cosmomc (July 2015 version)} package \citep{lewis02b}. 
CosmoMC invokes  CAMB\footnote{http://camb.info}  \citep{lewis00} to calculate the CMB power spectrum for each set of cosmological parameters. 
Although CosmoMC and CAMB have a partial implementation of PMF, we choose to adopt a simpler, fast template-based calculation for the PMF power. 
We have adapted CosmoMC to add a scaled version of the template to all four CMB power spectra: TT, TE, EE, and BB, where the scale factor is \apmf. 
\be
C_\ell = C_\ell (CAMB) + \apmf \cpmf
\ee
The calculation of the PMF template, \cpmf, is described later. 
Note that no other effect of the PMF is considered; this is one reason we only consider CMB data.

For the real data, unless noted, we assume the six parameter spatially flat \lcdm{} model with a single massive neutrino of 60\,meV and a seventh parameter \apmf{} describing the power due to the PMF. 
We adopt flat priors on all parameters. 
In particular, note that this is a flat prior on the observed PMF power, $\apmf$, not the rms magnetic field strength, \bpmf, that has often been used in the literature. 
In many chains, we also add \alens{} as a simple way of marginalizing over uncertainty in the predicted lensed BB power spectrum for any extension to the \lcdm{} model. 
Finally, in some cases, we allow non-zero tensors, parameterized as is normal, by the tensor-to-scalar ratio, r.

\subsection{Primordial Magnetic Field Template}

@@

\begin{figure*}[htb]\centering
\includegraphics[width=0.9\textwidth,clip,trim={1.5cm 12.cm 5cm 4cm}]{pmf_templates.pdf}
  \caption[CMB polarization from PMFs]{
  The expected contributions to CMB B-modes. 
  The red line shows the expected lensing B-mode power in the \planck{} best-fit \lcdm{} cosmology. 
  The dark green line corresponds to the contribution from inflationary gravitational waves for $r=0.05$; the solid black line is the sum of the inflationary gravitational wave and lensing B-modes. 
  The dashed purple line shows the PMF template used in this work, scaled to \apmf{} = 0.2. 
  As can be seen, the tensor contribution from PMFs at large angular scales (low $\ell$) closely resembles the expected inflationary gravitational wave signal. 
  However PMFs also produce a vector component that peaks around $\ell \sim 1700$ and allows the discrimination between  the inflationary and PMF explanations. 
      \label{fig:pmf-bb}
  }
\end{figure*}

@@

We find the addition of ground-based CMB polarization data to \planck{} data substantially improves the upper limits on the allowed power due to a PMF, leading to a factor of two improvement. 
With Planck alone, the 95\% confidence upper limit is $\apmf <  0.73$ for \lcdm{}+\apmf{}; with the ground-based experiments, this limit drops to $\apmf <  0.35$. 
The limits are slightly better with \alens{} allowed to vary (as a value for \alens{} slightly above unity is preferred). 
In this case for  Planck alone, the 95\% confidence upper limit is $\apmf <  0.68$, dropping to $\apmf <  0.27$ with the ground-based experiments. 
We test which datasets are important by adding single datasets to the \planck{} set while allowing \alens{} to float, and find that the majority of this improvement is due to the \bicepkeck{} ($\apmf <  0.28$) and \sptpol{} ($\apmf <  0.42$) B-mode power spectrum measurements. 
The other datasets do not have a large impact on the observed limits.

We propagate these limits on the observed PMF power into limits on the magnetic field strength, \bpmf. 
Due to the scaling with the fourth power, an upper limit with a flat prior on \apmf{} would lead to an apparent `detection' of \bpmf. 
We therefore importance sample the chains to generate a flat prior on \bpmf. 
We find a 95\% limit of $\bpmf < 1.97$\,nG for \planck{} alone, completely consistent with \citep{planck15-19} who found $\bpmf < 2.0$\,nG with $n_B=-2.9$ held fixed as we've done. 
This is slightly reduced to $\bpmf < 1.85$\,nG for the combined dataset. 
The improvement from adding more data is somewhat less than one might predict because the combined likelihood peaks about 1 sigma above zero, whereas the \planck{} only constraints peaks at zero.

\begin{table}[tbh]
\begin{center}
\caption{\label{tab:param_all} Parameter Constraints}
\small
\begin{tabular}{l | c c c }
Model   & $A_{pmf}$ &$A_{lens}$&$r$\\
\hline
\lcdm{} + \apmf{} &  $< 0.35$ & - & - \\
~~~+\alens{} & $< 0.27$ &$ 1.131 \pm  0.061$ &  -\\
~~~+\alens{}+r&  $< 0.24$&$ 1.133 \pm  0.062$ & $<0.056$\\ 
\end{tabular}
\tablecomments{ 
In the minimal model, \lcdm{} + \apmf{}, current CMB power spectra data from \planck, POLARBEAR, SPTpol, and \bicepkeck{} place a 95\% CL upper limit on the PMF power of $\apmf < 0.35$. 
This limit actually tightens slightly if tensor modes are allowed (r), or the amount of lensing is allowed to vary. 
In the case of r, the tightening is because the observed power at large angular scales is now split between PMFs or inflationary gravitational waves. 
In the case of \alens, it is because the preferred value for \alens{} is modestly above the predicted value of unity. 
Recall that the the PMF template is normalized to unity for \bpmf{} = 2.5\,nG.
} \normalsize
\end{center}
\end{table}

@@

The sensitivity of CMB experiments is increasing rapidly due to the exponential growth in detector counts
\citet{abazajian16} thus defines four stages of CMB experiments. 
The current experiments (i.e.~the ones used for PMF constraints in the last section) are classified as Stage II experiments. 
Stage III experiments such as SPT-3G, the Simons Array, or AdvACTPol \citep{benson14,suzuki16,henderson16} have approximately ten times more detectors, and generically will start collecting data in 2017 and finish in 2020-21.
In this section, we forecast the expected constraints from the Stage III experiments by combining forecasts for SPT-3G and the Simons Array. 
There is also a proposal to build a stage IV experiment, CMB-S4,  that would increase the detector counts by another order of magnitude and hopefully begin taking data in the early 2020s.
We examine the likely PMF constraints from CMB-S4, and consider which aspects of experimental or survey design are important to maximize the recovered information on PMFs.

@@

We use Fisher matrices to forecast the constraints on PMFs possible from each generation of experiment. 
Fisher matrices make it easy to combine different experiments and forecast the final parameter constraints. 
However, for non-Gaussian likelihoods, there is the potential for Fisher matrix forecasts to significantly underestimate the true uncertainties \citep[see e.g.,][]{wolz12}.

We include two external datasets in all forecasts. 
The first dataset is the expected measurements of the  TT, TE and EE spectra from the \planck{} satellite. 
We include \planck{} TT information in the multipole range $2\le \ell \le 3000$. 
Due to the importance of galactic foreground removal at large scales in polarization, we only use TE and EE information starting from $\ell = 30$. 
A prior on the optical depth of 0.005 is added to account for the expected optical depth constraint from the missing multipoles. 
Second, we include a 1\% external measurement of the Hubble constant, such as would be expected from the Taipan experiment \citep{kuehn14}.

One concern with combining Fisher matrices from different experiments is double-counting modes due to overlapping sky coverage. 
We sidestep this issue for the TT or TE spectra, by only using the \planck{} measurements. 
We do not expect the future experiments to substantially improve upon \planck{} in these spectra which are already cosmic variance limited out to fairly high multipoles. 
We throw away \planck{} EE data in the overlap region by appropriately increasing the forecasted \planck{} EE uncertainties. 
The exact increase depends on the assumed stage III or IV survey area.
We do not include \planck{} BB information in the forecasts. 
We ignore overlaps between the stage III  with the note that the overlap between a Chilean experiment like the Simons Array and South Pole experiment like SPT-3G should be small.
We do not include the stage III experiments, except as a prior on the polarized Poisson power, in the CMB-S4 constraints. 
With these measures, no modes should be double-counted in this analysis.

@@

~~~~SPT-3G & 6\% & 3.0 & 200 & 1.2 \\
~~~~Simons Array & 36\% & 9.5 & 200 & 3.5 \\ 
\\
%\hline
CMB Stage IV & 55\% & 1.3 & 100 & 4.0 \\
\end{tabular}
\tablecomments{ 
Key numbers about the planned stage III and IV experiments. 
The sky coverage percentages are after galactic cuts. 
Unless otherwise noted,  the Fisher matrix forecasts in this work use these numbers. 
All forecasts also include beam and calibration uncertainty as noted in the text. 
} \normalsize
\end{center}
\end{table*}

We list the assumed survey areas, noise levels and beam sizes for each experiment in Table~\ref{tab:experiments}. 
We also assume a 5\% uncertainty on the beam FWHM and a 1\% power calibration uncertainty; we have tested relaxing or tightening the beam and calibration uncertainty and find these calibration terms have a negligible impact on the PMF constraints.

We consider the constraints on PMFs in two cosmologies. 
In all cases, we also marginalize over unknown Poisson EE and BB  terms due to polarized extragalactic sources. 
Generally, we report constraints on the PMFs as one of several extensions to LCDM:  \lcdm{}  +r + \nrun{} + \neff{} + \mnu{}+ \apmf. 
However, in order to illustrate the degree to which parameter degeneracies are limiting the inferred constraints,  we also show constraints from an aggressive model in which the PMF power is the only extension to \lcdm{}:  \lcdm{}  + \apmf.

@@

The experiments that will begin taking data next year will dramatically improve constraints on the PMF power. 
With the minimal cosmological model, the 1-sigma forecasts for \planck{}+\ho{} is $\sigma(\apmf)=0.47$ as noted before.  
The uncertainty is forecast to fall nearly 400-fold by adding the two stage III experiments in this simplified model to $\sigma(\apmf)=0.0012$. 
Including all eleven parameters marginally weakens the \planck+\ho{} constraints on \apmf{} by 23\%, and the stage III constraint by a substantial factor of 2.5. 
Even so within this 11-parameter model, the addition of stage III CMB experiments improves the \apmf{} uncertainty by a factor of $\sim$\,180 to  $\sigma(\apmf)=0.0031$. 
We can expect substantially tighter constraints on PMFs by the end of the decade.

@@

The primary motivation behind the proposed CMB stage IV experiment is to search for inflationary gravitational waves and neutrino masses, however, the experiment should also enable extremely sensitive searches for PMFs. 
We begin by making forecasts for the fiducial CMB-S4 configuration as laid out in Table~\ref{tab:experiments}.

@@

Given that CMB-S4 is still being designed, it is worthwhile to consider how the design decisions might influence the final PMF constraints. 
We look at four aspects of the experiment design: the telescope size or beam FWHM, the low-$\ell$ noise performance (1/f knee), and the choice of survey area, and how well the beam size and calibration must be known. 
Of these four, we find that the low-$\ell$ noise performance and survey area are important to searches for PMFs

Larger telescopes will allow CMB polarization to be measured on smaller angular scales, and should therefor always improve the PMF constraints. 
However, larger telescopes are also more expensive to build which means at fixed cost, they would necessitate less ambitious focal planes and lower instantaneous mapping speed. 
We find the gains due to resolution at fixed mapping speed to be negligible: a beam size of FWHM=1$^\prime$ yields only a 1\% better limit than a FWHM of 10$^\prime$. 
This illustrates how important the large angular scales are to the PMF limits. 
We very crudely approximated a cost-neutral setup with a noise level of 1.3 \ukarcmin{} for a FWHM of 4$^\prime$; a noise level of 2.9 \ukarcmin{} for a FWHM of 2$^\prime$; and a noise level of 4.1 \ukarcmin{} for a FWHM of 1$^\prime$. 
In this quasi-cost-neutral setup, there is a strong preference to go for more detectors over better angular resolution: the uncertainty was a factor of 2.5 times smaller in the coarse case over the FWHM=1$^\prime$ case. 
Raw sensitivity is more important than angular resolution to an experiment's ability to study PMFs.

@@

In this work, we have improved the current upper limits on the strength of any primordial magnetic fields by including more CMB polarization data. 
By adding \bicepkeck{}, \pb, and \sptpol to \planck{} we find the 95\% CL upper limit on the PMF power falls from $\apmf < 0.68$ to $\apmf < 0.33$. 
The biggest contributor to this improvement is the low-$\ell$ data from \bicepkeck{}.

We have also shown that the next generation of experiments have the potential power to dramatically reduce these limits, and potentially detect PMFs for the first time. 
The so-called stage III experiments, which should begin taking data in 2017, expect to set upper limits at the level of $\apmf < 0.0062$ even after marginalizing over a five-parameter extension to \lcdm.

>>>>

\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\planck}{{\sl Planck}}
\newcommand{\wmap}{{\sl WMAP}}
\newcommand{\bicepkeck}{BICEP2/Keck Array}
\newcommand{\sptnew}{SPT-3G}
\newcommand{\pb}{POLARBEAR}
\newcommand{\simons}{Simons Array}
\newcommand{\sptpol}{SPTpol}
\newcommand{\advactpol}{Adv.~ACTpol}

@@

We present new limits on the amplitude of potential primordial magnetic fields (PMFs) using temperature and polarization measurements of the cosmic microwave background (CMB)  from \planck{}, \bicepkeck{}, \pb, and \sptpol. 
We reduce twofold the upper limit on the CMB anisotropy power due to a PMF, from $\apmf < 0.73$ from Planck alone to $\apmf < 0.35$ for the combined dataset at 95\% CL. 
%Most of the improvement is due to the addition of the \bicepkeck{} data; without these bandpowers, the combined limit weakens to $\apmf < 0.\tbd{XX}$. 
We also forecast the expected future limits from soon-to-deploy CMB experiments (like \sptnew{},  \advactpol, or the \simons) and the proposed CMB-S4 experiment. 
Future CMB experiments should dramatically reduce the current uncertainties, by two orders of magnitude for the near-term experiments and three orders of magnitude for the CMB-S4 experiment. 
The constraints from CMB-S4 have the potential to rule out much of the parameter space for PMFs.

@@

CMB B-mode measurements can also be used to constrain more exotic models, such as the possible existence of cosmic birefringence \citep{carroll98,lue99} or a primordial magnetic field (PMF) \citep{kosowsky96, seshadri01}.  
Both effects lead to cosmic birefringence and the rotation of E-modes into B-modes. 
By equating the magnitude of the resulting B-modes, parity-violating processes can be translated into an equivalent PMF strength. 
We will therefore simply quote effective limits on a PMF. 
In this work, we present new upper limits from current CMB polarization data on the possibility of a PMF or parity-violating physics.

@@

PMFs would have observational consequences for Big Bang nucleosynthesis \citep[e.g.,][]{kahniashvili10}, large scale structure \citep[e.g.,][]{battaner97}, the ionization history of the Universe \citep{kunze15}, and the black body spectrum of the CMB \citep[e.g.,][]{kunze14},  as well as the CMB anisotropies. 
The CMB anisotropies have yielded some of the strongest constraints on PMFs and are the focus of this work.
There have been two recent results of note. 
\citet{planck15-19} have used the \planck{} 2015 release of temperature and polarization data to set limits on a variety of PMF models. 
With the CMB power spectrum data that will be the focus of this paper, \citet{planck15-19} find 95\% CL upper limits ranging from $\bpmf < 5.6$\,nG to $<0.7$\,nG depending on the exact model. 
The \pb{} collaboration also recently announced limits on PMFs from the \pb{} data using either a 4-point estimator or B-mode power spectrum measurement in \citep{polarbear14b}. 
The strongest constraints were from the B-mode spectrum; the observed upper limit was $\bpmf < 3.9$\,nG. 
Two other experiments, \sptpol{} and \bicepkeck{}, have reported B-mode power spectrum measurements recently \citep{keisler15,bicepkeck15}.
In this paper, we will combine the data from all four experiments to determine simple limits on the PMF, and then examine Fisher matrix forecasts on PMF models from the stage-III and stage-IV CMB experiments being built or designed right now.

@@

In addition to the \planck{} data, we use a number of recent CMB B-mode polarization measurements. 
First, we include measurements of the TE, EE and BB power spectra from the \sptpol{} experiment \citep{crites15,keisler15}. 
The BB bandpowers cover the angular multipoles $\ell \in [500,2000]$. 
We also add the BB bandpowers from \pb{} that cover the multipoles from 500 to 2500 \citep{polarbear14b}. 
At these angular scales, both the \sptpol{} and \pb{} bandpowers primary constrain the vector modes of a PMF. 
Finally, we include the latest BICEP2 and Keck Array  joint analysis \citep{bicepkeck15}. 
This latter dataset also places limits on the tensor modes of a PMF due to its coverage of lower multipoles. 
During the writing of this work, ACTpol polarization power spectra became available \citep{naess14,louis16}. 
We have not included them as the public likelihoods do not include the B-mode power spectrum; we do not expect adding the ACTpol bandpowers to significantly change the PMF limits based on a visual comparison of the current BB bandpowers from different experiments.
In all chains, we marginalize over the recommended foreground models for each data set. 
We do not however require these consistency between these foreground models since the data don't have identical flux cuts for extragalactic sources and so on. 
In the case of an upper limit this treatment is conservative since wider foreground priors should only weaken the limit on the PMF strength.

@@

We use MCMC methods to determine the parameter constraints reported in this work. 
The results are calculated using  the {\textsc CosmoMC}\footnote{http://cosmologist.info/cosmomc (July 2015 version)} package \citep{lewis02b}. 
CosmoMC invokes  CAMB\footnote{http://camb.info}  \citep{lewis00} to calculate the CMB power spectrum for each set of cosmological parameters. 
Although CosmoMC and CAMB have a partial implementation of PMF, we choose to adopt a simpler, fast template-based calculation for the PMF power. 
We have adapted CosmoMC to add a scaled version of the template to all four CMB power spectra: TT, TE, EE, and BB, where the scale factor is \apmf. 
\be
C_\ell = C_\ell (CAMB) + \apmf \cpmf
\ee
The calculation of the PMF template, \cpmf, is described in \S\ref{sec:template}. 
Note that no other effect of the PMF is considered; this is one reason we only consider CMB data.

For the real data, unless noted, we assume the six parameter spatially flat \lcdm{} model with a single massive neutrino of 60\,meV and a seventh parameter \apmf{} describing the power due to the PMF. 
We adopt flat priors on all parameters. 
In particular, note that this is a flat prior on the observed PMF power, $\apmf$, not the rms magnetic field strength, \bpmf, that has often been used in the literature. 
In many chains, we also add \alens{} as a simple way of marginalizing over uncertainty in the predicted lensed BB power spectrum for any extension to the \lcdm{} model. 
Finally, in some cases, we allow non-zero tensors, parameterized by the tensor-to-scalar ratio, r.

\subsection{Primordial Magnetic Field Template}
\label{sec:template}

@@

\begin{figure*}[htb]\centering
\includegraphics[width=0.9\textwidth,clip,trim={1.5cm 12.cm 5cm 4cm}]{pmf_templates.pdf}
  \caption[CMB polarization from PMFs]{
  The expected contributions to CMB B-modes. 
  The red line shows the expected lensing B-mode power in the \planck{} best-fit \lcdm{} cosmology. 
  The dark green line corresponds to the contribution from inflationary gravitational waves for $r=0.05$; the solid black line is the sum of the inflationary gravitational wave and lensing B-modes. 
  The dashed purple line shows the PMF template used in this work, scaled to \apmf{} = 0.2 to match the gravitational wave signal. 
  As can be seen, the tensor contribution from PMFs at large angular scales (low $\ell$) closely resembles the expected inflationary gravitational wave signal, and the two signals are effectively degenerate in a power spectrum (2-point) analysis.
  However, if excess power on these scales is detected, higher order moments of the map (i.e.~4-point estimators) might be used to discriminate between   the inflationary and PMF explanations. 
      \label{fig:pmf-bb}
  }
\end{figure*}

@@

We find the addition of ground-based CMB polarization data to \planck{} data substantially improves the upper limits on the allowed power due to a PMF, leading to a factor of two improvement. 
With Planck alone, the 95\% confidence upper limit is $\apmf <  0.73$ for \lcdm{}+\apmf{}; with the ground-based experiments, this limit drops to $\apmf <  0.35$. 
Somewhat counterintuitively, the limits are slightly better with \alens{} allowed to vary (this is due to a preference for \alens{} above unity). 
In this case for  Planck alone, the 95\% confidence upper limit is $\apmf <  0.68$, dropping to $\apmf <  0.27$ with the ground-based experiments. 
We test which datasets are important by adding single datasets to the \planck{} set while allowing \alens{} to float, and find that the majority of this improvement is due to the \bicepkeck{} ($\apmf <  0.28$) and \sptpol{} ($\apmf <  0.42$) B-mode power spectrum measurements. 
The other datasets do not have a large impact on the observed limits.

We propagate these limits on the observed PMF power into limits on the magnetic field strength, \bpmf. 
Due to the scaling with the fourth power, an upper limit with a flat prior on \apmf{} would lead to an apparent `detection' of \bpmf. 
We thereforee importance sample the chains to generate a flat prior on \bpmf. 
We find a 95\% limit of $\bpmf < 1.97$\,nG for \planck{} alone, completely consistent with \citep{planck15-19} who found $\bpmf < 2.0$\,nG with $n_B=-2.9$ held fixed as we've done. 
This is slightly reduced to $\bpmf < 1.85$\,nG for the combined dataset. 
The improvement from adding more data is somewhat less than one might predict because the combined likelihood peaks somewhat above zero, whereas the \planck{}-only constraints peaks at zero.

\begin{table}[tbh]
\begin{center}
\caption{\label{tab:param_all} Parameter Constraints}
\small
\begin{tabular}{l | c c c }
Model   & $A_{pmf}$ &$A_{lens}$&$r$\\
\hline
\lcdm{} + \apmf{} &  $< 0.35$ & - & - \\
~~~+\alens{} & $< 0.27$ &$ 1.131 \pm  0.061$ &  -\\
~~~+\alens{}+r&  $< 0.24$&$ 1.133 \pm  0.062$ & $<0.056$\\ 
\end{tabular}
\tablecomments{ 
In the minimal model, \lcdm{} + \apmf{}, current CMB power spectra data from \planck, POLARBEAR, SPTpol, and \bicepkeck{} place a 95\% CL upper limit on the PMF power of $\apmf < 0.35$. 
This limit tightens slightly if tensor modes are allowed (r), or the amount of lensing is allowed to vary. 
In the case of r, the tightening is because the observed power at large angular scales is now split between PMFs or inflationary gravitational waves. 
In the case of \alens, it is because the preferred value for \alens{} is modestly above the predicted value of unity. 
Recall that the the PMF template is normalized to unity for \bpmf{} = 2.5\,nG.
} \normalsize
\end{center}
\end{table}

@@

The sensitivity of CMB experiments is increasing rapidly due to the exponential growth in detector counts. 
\citet{abazajian16} thus defines four stages of CMB experiments. 
The current experiments (i.e.~the ones used for PMF constraints in the last section) are classified as Stage II experiments. 
Stage III experiments such as SPT-3G, the Simons Array, or AdvACTPol \citep{benson14,suzuki16,henderson16} have approximately ten times more detectors, and generically will start collecting data in 2017 and finish in 2020-21.
In this section, we forecast the expected constraints from the Stage III experiments by combining forecasts for SPT-3G and the Simons Array. 
There is also a proposal to build a stage IV experiment, CMB-S4,  that would increase the detector counts by another order of magnitude and hopefully begin taking data in the early 2020s.
We examine the likely PMF constraints from CMB-S4, and consider which aspects of experimental or survey design are important to maximize the recovered information on PMFs.

@@

We use Fisher matrices to forecast the constraints on PMFs possible from each generation of experiment. 
While Fisher matrices may underestimate the true uncertainties for non-Gaussian likelihoods \citep[see e.g.,][]{wolz12}, they 
 make it easy to combine different experiments and estimate the final parameter constraints.

We include two external datasets in all forecasts. 
The first dataset is the expected measurements of the  TT, TE and EE spectra from the \planck{} satellite. 
We include \planck{} TT information in the multipole range $2\le \ell \le 3000$. 
Due to the importance of galactic foreground removal at large scales in polarization, we only use TE and EE information starting from $\ell = 30$. 
A prior on the optical depth of 0.005 is added to account for the expected optical depth constraint from the missing multipoles. 
Second, we include a 1\% external measurement of the Hubble constant, such as is expected from the Taipan experiment \citep{kuehn14}.

One concern with combining Fisher matrices from different experiments is double-counting modes due to overlapping sky coverage. 
We avoid this problem in different ways in each of the CMB power spectra. 
For the TT and TE spectra, we avoid double-counting by only using the \planck{} measurements. 
We do not expect the future experiments to substantially improve upon \planck{} in these spectra which are already cosmic variance limited out to fairly high multipoles. 
We do the opposite for the EE and BB spectra. 
We do not include \planck{} BB information in any forecast, and  we throw away \planck{} EE data in the overlap region by appropriately increasing the forecasted \planck{} EE uncertainties. 
%The exact increase depends on the assumed stage III or IV survey area.
We ignore overlaps between the stage III  with the note that the sky overlap between a Chilean experiment like the Simons Array and South Pole experiment like SPT-3G should be small. 
The overlapping sky coverage will probably be substantial between stage III and IV experiments, thus we do not include the stage III experiments, except as a prior on the polarized Poisson power, in the CMB-S4 constraints. 
With these measures, no modes should be double-counted in this analysis.

@@

~~~~SPT-3G & 6\% & 3.0 & 200 & 1.2 \\
~~~~Simons Array & 36\% & 9.5 & 200 & 3.5 \\ 
\\
%\hline
CMB Stage IV & 55\% & 1.3 & 100 & 4.0 \\
\end{tabular}
\tablecomments{ 
Key numbers about the planned stage III and IV experiments. 
The sky coverage percentages are after galactic cuts. 
Unless otherwise noted,  the Fisher matrix forecasts in this work use these numbers. 
All forecasts also include beam and calibration uncertainties as noted in the text. 
} \normalsize
\end{center}
\end{table*}

We list the assumed survey areas, noise levels and beam sizes for each experiment in Table~\ref{tab:experiments}. 
We  assume a 5\% uncertainty on the beam FWHM and a 1\% power calibration uncertainty; we have tested relaxing or tightening the beam and calibration uncertainty and find these calibration terms have a negligible impact on the PMF constraints.

We consider the constraints on PMFs in two cosmologies. 
The first is a 11-parameter model that extends \lcdm{} with 5 commonly considered extensions:  \lcdm{}  +r + \nrun{} + \neff{} + \mnu{}+ \apmf. 
This 11-parameter model is our default cosmological model for forecasting PMF constraints. 
Additionally, in order to illustrate the degree to which parameter degeneracies are limiting the inferred constraints,  we show constraints from a `minimal' 7-parameter model in which the PMF power is the only extension to \lcdm{}:  \lcdm{}  + \apmf. 
In both cases, we also marginalize over unknown Poisson EE and BB  terms due to polarized extragalactic sources.

In this work, we only forecast the uncertainties due to power spectrum (i.e.~2-point estimators) searches for PMFs. 
Currently the power spectrum limits from \planck{} or \pb{} are better than the 4-point upper limits. 
Although outside the scope of this work, it would be interesting to extend this analysis to 4-point estimators for two reasons. 
First, the 4-point limits should improve faster as the noise level falls, and second, in the case of a detection, the 4-point estimators might discriminate between inflationary gravitational waves and tensor PMFs.

@@

The experiments that will begin taking data in 2017 will dramatically improve constraints on the PMF power. 
With the minimal cosmological model, the 1-sigma forecasts for \planck{}+\ho{} is $\sigma(\apmf)=0.47$ (this is consistent with, although slightly worse than, the upper limits from the real data).  
Adding the two stage III experiments (\sptnew{} and \simons{}) reduces the uncertainty nearly 400-fold in this 7-parameter model to $\sigma(\apmf)=0.0012$. 
Including all eleven parameters marginally weakens the \planck+\ho{} constraints on \apmf{} by 23\%, and the stage III constraint by a substantial factor of 2.5. 
Even within the 11-parameter model however,  the addition of stage III CMB experiments improves the \apmf{} uncertainty by a factor of $\sim$\,180 to  $\sigma(\apmf)=0.0031$. 
We can expect substantially tighter constraints on PMFs by the end of the decade.

@@

The primary motivation behind the proposed CMB stage IV experiment commonly called CMB-S4 is to search for inflationary gravitational waves and neutrino masses, however, the experiment should also enable extremely sensitive searches for PMFs. 
We begin by making forecasts for the fiducial CMB-S4 configuration as laid out in Table~\ref{tab:experiments}.

@@

Given that CMB-S4 is still being designed, it is worthwhile to consider how design decisions will affect the final PMF constraints. 
We look at four aspects of the experiment design: the telescope size or beam FWHM, the low-$\ell$ noise performance (1/f knee),  the choice of survey area, and how well the beam size and calibration must be known. 
Of these four, we find that the low-$\ell$ noise performance and survey area are important to searches for PMFs

Larger telescopes will allow CMB polarization to be measured on smaller angular scales, and should therefore always improve the PMF constraints. 
However, larger telescopes are also more expensive to build which means at fixed cost, they would necessitate less ambitious focal planes and lower instantaneous mapping speed. 
We find the gains due to resolution at fixed mapping speed to be negligible: a beam size of FWHM=1$^\prime$ yields only a 1\% better limit than a FWHM of 10$^\prime$. 
This illustrates how important the large angular scales are to the PMF limits. 
We very crudely approximated a cost-neutral setup with a noise level of 1.3 \ukarcmin{} for a FWHM of 4$^\prime$; a noise level of 2.9 \ukarcmin{} for a FWHM of 2$^\prime$; and a noise level of 4.1 \ukarcmin{} for a FWHM of 1$^\prime$. 
In this quasi-cost-neutral setup, there is a strong preference to go for more detectors over better angular resolution: the uncertainty was a factor of 2.5 times smaller in the coarse case over the FWHM=1$^\prime$ case. 
Raw sensitivity is more important than angular resolution to an experiment's ability to study PMFs.

@@

In this work, we have improved the current upper limits on the strength of any primordial magnetic fields by including more CMB B-mode polarization data. 
By adding \bicepkeck{}, \pb, and \sptpol{} to \planck{} we find the 95\% CL upper limit on the PMF power falls from $\apmf < 0.73$ to $\apmf < 0.35$. 
The biggest contributor to this improvement is the low-$\ell$ data from \bicepkeck{}.

We have also shown that the next generation of experiments should dramatically reduce these limits, and potentially detect PMFs for the first time. 
These so-called stage III experiments, which will begin taking data in 2017, expect to set upper limits at the level of $\apmf < 0.0062$ even after marginalizing over a five-parameter extension to \lcdm.
